---
title: "Homework 4"
author: "Drew Dahlquist"
date: "3/1/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

$$\\[3in]$$

### 2.

(a)

P(Y = 1 | X) = `r 1-1/(1.12)`

(b)

P(Y = 1 | X) = `r 0.52/0.48`

### 3.

```{r 3}
Auto = read.csv("Auto.csv", na.strings="?")
Auto = na.omit(Auto)
Auto.class = Auto[,c(1,2,4,6,7,8)]
Auto.class = Auto.class[Auto.class$origin != '2',]
Auto.class$origin = as.factor(Auto.class$origin)
summary(Auto.class)
```

(a)

```{r 3a}
glm.fit = glm(origin ~ mpg + cylinders + horsepower + acceleration + year, family=binomial, data=Auto.class)
summary(glm.fit)
```

Yes, mpg, cylinders, and year appear to be statistically significant according to the summary.

(b)

The estimated coefficient for cylinders indicates that a 1-unit increase in cylinders will decrease the log-odds of origin by -1.19. Similarly, the coefficient for year indicates that a 1-unit increase in year will decrease the log-odds of origin by -0.19.

(c)

```{r 3c}
glm.probs = predict(glm.fit,type="response")
contrasts(Auto.class$origin)
glm.pred=rep("1", length(Auto.class$origin))
glm.pred[glm.probs > 0.5] = "3"
table(glm.pred, Auto.class$origin)
```

Total fraction of correct predictions is `r (219+57)/length(Auto.class$origin)`.

(d)

```{r 3d}
library(MASS)
lda.fit=lda(origin ~ mpg + cylinders + horsepower + acceleration + year, data=Auto.class)
lda.fit
```

(e)

```{r 3e}
plot(lda.fit)
```

(f)

```{r 3f}
lda.pred=predict(lda.fit, Auto.class)
lda.class=lda.pred$class
table(lda.class, Auto.class$origin)
```

(g)

```{r 3g}
qda.fit=qda(origin ~ mpg + cylinders + horsepower + acceleration + year, data=Auto.class)
qda.fit
```

(h)

```{r 3h}
qda.pred=predict(qda.fit, Auto.class)
qda.class=qda.pred$class
table(qda.class, Auto.class$origin)
```

(i)

Comparison of LDA vs QDA

|             |    LDA           |   QDA            |
|-------------|------------------|------------------|
| Accuracy    | `r (220+56)/324` | `r (194+69)/324` |
| Sensitivity | `r 56/(25+56)`   | `r 69/(51+69)`   |
| Specificity | `r 23/(220+23)`  | `r 10/(194+10)`  |
| Precision   | `r 56/(23+56)`   | `r 69/(10+69)`   |
