---
title: "STAT 4510 HW1"
author: "Drew Dahlquist"
date: "1/21/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### 2.3.1 Basic Commands

```{r}
x = c(1,2,3,5)
x
```

```{r}
x = c(1,6,2)
x
y = c(1,4,3)
```

```{r}
length(x)
length(y)
x+y
```

```{r}
ls()
rm(x,y)
ls()
```

```{r}
rm(list=ls())
```

```{r}
?matrix
```

```{r}
x=matrix(data=c(1,2,3,4), nrow=2, ncol=2)
x
```

```{r}
matrix(c(1,2,3,4),2,2,byrow=TRUE)
```

```{r}
sqrt(x)
x^2
```

```{r}
x=rnorm(50)
y=x+rnorm(50,mean=50,sd=.1)
cor(x,y)
```

```{r}
set.seed(1303)
rnorm(50)
```

```{r}
set.seed(3)
y=rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)
```

### 2.3.2 Graphics

```{r}
x=rnorm(100)
y=rnorm(100)
plot(x,y)
plot(x,y,xlab="this is the x-axis",ylab="this is the y-axis", main="Plot of X vs Y")
```

```{r}
pdf("Figure.pdf")
plot(x,y,,col="green")
dev.off()
```

```{r}
x=seq(1,10)
x
x=1:10
x
x=seq(-pi,pi,length=50)
```

```{r}
y=x
f=outer(x,y,function(x,y)cos(y)/(1+x^2))
contour(x,y,f)
contour(x,y,f,nlevels=45,add=T)
fa=(f-t(f))/2
contour(x,y,fa,nlevels=15)
```

```{r}
image(x,y,fa)
persp(x,y,fa)
persp(x,y,fa,theta=30)
persp(x,y,fa,theta=30,phi=20)
persp(x,y,fa,theta=30,phi=70)
persp(x,y,fa,theta=30,phi=40)
```

### 2.3.3 Indexing Data

```{r}
A=matrix(1:16,4,4)
A
```

```{r}
A[2,3]
```

```{r}
A[c(1,3),c(2,4)]
A[1:3,2:4]
A[1:2,]
A[,1:2]
```

```{r}
A[1,]
```

```{r}
A[-c(1,3),]
A[-c(1,3),-c(1,3,4)]
```

```{r}
dim(A)
```

### 2.3.4 Loading Data

```{r}
Auto=read.table("Auto.data")
fix(Auto)
```

```{r}
Auto=read.table("Auto.data",header=T,na.strings="?")
fix(Auto)
```

```{r}
Auto=read.csv("Auto.csv",header=T,na.strings="?")
fix(Auto)
dim(Auto)
Auto[1:4,]
```

```{r}
Auto=na.omit(Auto)
dim(Auto)
```

```{r}
names(Auto)
```

### 2.3.5 Additional Graphical and Numerical Summaries

```{}
plot(cylinders, mpg)
```

```{r}
plot(Auto$cylinders, Auto$mpg)
attach(Auto)
plot(cylinders, mpg)
```

```{r}
cylinders=as.factor(cylinders)
```

```{r}
plot(cylinders, mpg)
plot(cylinders, mpg, col="red")
plot(cylinders, mpg, col="red", varwidth=T)
plot(cylinders, mpg, col="red", varwidth=T, horizontal=T)
plot(cylinders, mpg, col="red", varwidth=T, xlab="cylinders", ylab="MPG")
```

```{r}
hist(mpg)
hist(mpg, col=2)
hist(mpg, col=2, breaks=15)
```

```{}
pairs(Auto)
```
```{r}
pairs(~ mpg + displacement + horsepower + weight + acceleration, Auto)
```

```{r}
plot(Auto$horsepower, mpg)
identify(horsepower, mpg, name)
```

```{r}
summary(Auto)
```

```{r}
summary(mpg)
```

## Problem 2

```{r echo=FALSE}
rm(list=ls())
```

### a)

```{r}
?rnorm
```
The `rnorm` function generates a normal distribution with given mean and sd. You may also give `rnorm` a vector of quantiles or probabilities, a number of observations to make. It also allows for given probabilites to be given as log(p) or as P[X < x].

### b)

```{r echo=FALSE}
x = 8.5
red_err = abs(5+0.25*x^2 - (4.8+0.27*x^2))
```

The reducible error for the prediction is `r red_err`.

### c)

```{r echo=FALSE}
# observation
set.seed(45107510)
x <- rnorm(n=1000, mean=4, sd=2) # locations of x
ep <- rnorm(n=1000, mean=0, sd=2) # error terms
y <- 5+0.25*x^2+ep
p1 <- plot(x, y, xlab="", ylab="", col="blue")
# true underlying function
xGrid <- seq(-2, 10, by=0.1)
fTrue <- 5+0.25*xGrid^2
p2 <- lines(xGrid, fTrue, col="orange", lwd=3)
# plus minus two standard deviation
plusTwoSd <- fTrue + 2
minusTwoSd <- fTrue - 2
p3 <- lines(xGrid, plusTwoSd, col="orange", lwd=3, lty=2)
p4 <- lines(xGrid, minusTwoSd, col="orange", lwd=3, lty=2)
# legend
legend(-2, 28, legend=c(expression(f(X)), expression(f(X) %+-% 2)),col=c("orange", "orange"), lty=1:2, lwd=c(3,3))
```

After changing the sd of the error terms to 2, the sampled data points become less distributed around the true underlying function.

## Problem 3

```{r echo=FALSE}
rm(list=ls())
```

### a)

`install.packages("ISLR")`
```{r}
library(ISLR)

?Credit
```

### b)

```{r}
credit = read.csv("Credit.csv")
```

### c)

```{r}
plot(credit$Income, credit$Balance)
plot(credit$Age, credit$Balance)
plot(credit$Limit, credit$Balance)
```

### d)

```{r}
summary(credit)
credit$Own = as.factor(credit$Own)
credit$Student = as.factor(credit$Student)
credit$Married = as.factor(credit$Married)
credit$Region = as.factor(credit$Region)
```

### e)

```{r}
summary(credit)
```

The factor variables now the amount of observations for each category for the variable.

### f)

```{r}
hist(credit$Income, main="Histogram of Income", xlab="Income", ylab="Frequency", col="2")
hist(credit$Age, main="Histogram of Age", xlab="Age", ylab="Frequency", col="3")
```

### g)

```{r}
boxplot(credit$Limit, main="Boxplot of Limit", ylab="Limit", col="2")
boxplot(credit$Balance, main="Boxplot of Balance", ylab="Balance", col="3")
```
